taxize: taxonomic search and retrieval in R

### 2013-10-01 f1000_rev/taxize_appB.Rnw

\section{Matching species tables with different taxonomic resolution}

@@

This is just an example how taxonomic APIs (via taxize) could be used to search for matches up- and downwards the taxonomic ladder. We are looking forward to integrate the freshwaterecology.info database \url{www.freshwaterecology.info} into taxize, which will facilitate trait-based analyses in R.

>>>>

\documentclass[letterpaper,superscriptaddress,showkeys,longbibliography]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
    colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

\begin{document}
<<setup, include=FALSE, cache=FALSE, size='scriptsize'>>=
## set global chunk options
opts_chunk$set(fig.align='center', fig.show='hold', par=TRUE, warning=FALSE, message=FALSE, error=FALSE, comment='#')
@

\section*{Appendix B. Matching species tables with different taxonomic resolution}

@@

This is just an example how taxonomic APIs (via taxize) could be used to search for matches up- and downwards the taxonomic ladder. We are looking forward to integrate the freshwaterecology.info database \url{www.freshwaterecology.info} into taxize, which will facilitate trait-based analyses in R.
\end{document}

### 2013-10-01 f1000_rev/taxize_ms.tex

A user can use the mentioned \emph{classification}-function to retrieve the taxonomic hierarchy and then search the hierarchies up- and downwards for matches. Here is an example to match a species with names on three different taxonomic levels.

\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{A} \hlkwb{<-} \hlstr{"gammarus roeseli"}
\hlstd{B1} \hlkwb{<-} \hlstr{"gammarus roeseli"}
\hlstd{B2} \hlkwb{<-} \hlstr{"gammarus"}
\hlstd{B3} \hlkwb{<-} \hlstr{"gammaridae"}
\hlstd{A_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(A,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B1_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(B1,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B2_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(B2,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B3_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(B3,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B1[}\hlkwd{match}\hlstd{(A, B1)]}
\end{alltt}
\begin{verbatim}
# [1] "gammarus roeseli"
\end{verbatim}
\begin{alltt}
\hlstd{A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{Rank[}\hlkwd{tolower}\hlstd{(A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{ScientificName)} \hlopt{%in%} \hlstd{B2]}
\end{alltt}
\begin{verbatim}
# [1] "genus"
\end{verbatim}
\begin{alltt}
\hlstd{A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{Rank[}\hlkwd{tolower}\hlstd{(A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{ScientificName)} \hlopt{%in%} \hlstd{B3]}
\end{alltt}
\begin{verbatim}
# [1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

If one finds a direct match (here \emph{Gammarus roeseli}), they will be lucky. But you can also match Gammaridae with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait table with an abundance table) is given in Appendix B.

>>>>

A user can use the mentioned \emph{classification}-function to retrieve the taxonomic hierarchy and then search the hierarchies up- and downwards for matches. Here is an example to match a species (A) with names of on different taxonomic levels (B1 \& B2).

\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{A} \hlkwb{<-} \hlstr{"gammarus roeseli"}
\hlstd{B1} \hlkwb{<-} \hlstr{"gammarus"}
\hlstd{B2} \hlkwb{<-} \hlstr{"gammaridae"}
\hlstd{A_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(A,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B1_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(B1,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{B2_clas} \hlkwb{<-} \hlkwd{classification}\hlstd{(B2,} \hlkwc{db} \hlstd{=} \hlstr{'ncbi'}\hlstd{)}
\hlstd{A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{Rank[}\hlkwd{tolower}\hlstd{(A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{ScientificName)} \hlopt{%in%} \hlstd{B1]}
\end{alltt}
\begin{verbatim}
# [1] "genus"
\end{verbatim}
\begin{alltt}
\hlstd{A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{Rank[}\hlkwd{tolower}\hlstd{(A_clas[[}\hlnum{1}\hlstd{]]}\hlopt{$}\hlstd{ScientificName)} \hlopt{%in%} \hlstd{B2]}
\end{alltt}
\begin{verbatim}
# [1] "family"
\end{verbatim}
\end{kframe}
\end{knitrout}

If one finds a direct match (here \emph{Gammarus roeseli}), they will be lucky. However, Gammaridae can also be matched with \emph{Gammarus roeseli}, but on a lower taxonomic level. A more comprehensive and realistic example (matching a trait table with an abundance table) is given in Appendix B.
